|

### 2019-12-30 inst/manuscript/candidaev_supplement.Rmd
Add Supplementary Figure S3
Boxplots comparing mode EV size and EV particle
concentration for each 'strain'.

# load libraries
library(candidaev)
library(tidyNano)
library(limma)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(kableExtra)
```

@@

```{r tableS1Data}
# construct data.frame with nta experiment info
nta_metadata <- data.frame(id = c("ATCC10231_1", "ATCC10231_2", "ATCC10231_3", 
                                  "ATCC90028_1", "ATCC90028_2", "ATCC90028_3", 
                                  "Biofilm_1", "Biofilm_2", "Biofilm_3", 
                                  "Biofilm_4", "Biofilm_5", 
                                  "DAY286_1", "DAY286_2", "DAY286_3"), 
                           strain = c(rep("ATCC10231", 3), 
                                      rep("ATCC90028", 3), 
                                      rep("Biofilm", 5), 
                                      rep("DAY286", 3)), 
                           replicate = c(seq(1, 3), seq(1, 3), 
                                         seq(1, 5), seq(1, 3)), 
                           OD600 = c(0.455*20, 0.557*20, 0.615*20, 
                                     0.310*50, 0.337*50, 0.297*50, 
                                     rep(NA, 5), 
                                     36.2, 39.7, 34.0),
                           culture_vol = c(rep(150, 3),
                                           rep(150, 3), 
                                           rep(1000, 5), 
                                           rep(300, 3)), 
                           protein_yield = c(19.2, 28.2, 28.7, 
                                             42.4, 71.6, 58.4, 
                                             20.4, 19.8, 29.7, 19.3, 41.6, 
                                             21.9, 18.0, 22.4), 
                           protein_conc = c(0.192, 0.282, 0.287, 
                                            0.424, 0.716, 0.584, 
                                            0.204, 0.198, 0.297, 0.193, 0.416, 
                                            0.437, 0.360, 0.447), 
                           df = c(250, 750, 750, 
                                  1500, 3000, 3000, 
                                  3000, 2000, 2000, 3000, 6000, 
                                  1500, 1500, 1500), 
                           mean = c(248.5, 218.7, 217.9, 
                                    197.1, 233.1, 191.0, 
                                    119.4, 142.1, 143.3, 133.3, 120.9, 
                                    182.8, 174.5, 164.5), 
                           mode = c(171.9, 155.5, 140.2, 
                                    135.4, 156.7, 138.0, 
                                    100.5, 94.7, 103.1, 103.1, 97.5, 
                                    106.4, 99.2, 100.2), 
                           sd = c(120.1, 112.1, 110.0, 
                                  87.1, 90.7, 68.4, 
                                  47.4, 64.3, 70.2, 60.7, 47.4, 
                                  93.4, 93.9, 83.1), 
                           d10 = c(139.1, 110.2, 118.1, 
                                   122.4, 141.9, 126.0, 
                                   90.2, 88.8, 96.4, 93.2, 89.7, 
                                   97.8, 93.8, 93.0), 
                           d50 = c(211.9, 188.2, 180.3, 
                                   169.4, 209.3, 170.7, 
                                   105.6, 115.0, 110.9, 108.1, 105.9, 
                                   152.8, 147.3, 136.2),
                           d90 = c(393.0, 354, 354.4, 
                                   301.4, 381.1, 285.8, 
                                   151.5, 239.7, 240.8, 210.2, 171.2, 
                                   320.4, 308.0, 290.9),
                           sort_order = c(rep(3, 3), rep(2, 3), 
                                          rep(4, 5), rep(1, 3)), 
                           stringsAsFactors = FALSE)

# determine the total particle concentration for each biological replicate
nta_total <- nta_techavg %>% 
  nanocount(Strain, Bio_rep, Dilution, 
            param_var = Tech_rep_mean) %>% 
  tidyr::unite("id", Strain, Bio_rep) %>% 
  inner_join(., nta_metadata, by = "id") %>% 
  arrange(sort_order) %>% 
  select(strain, replicate, OD600, culture_vol, protein_yield, protein_conc, 
         Particle_count, mean, mode, sd) %>% 
  mutate(OD600 = case_when(is.na(OD600) == TRUE ~ "NA", 
                           TRUE ~ as.character(OD600)), 
         Particle_count = signif(Particle_count, digits = 3))
```

@@

# construct a table of metadata for each EV biological replicate
tabS1 <- nta_total %>% 
  select(-strain) %>% 
  mutate_all(linebreak) %>% 
  knitr::kable(caption = tabS1_cap, 
               format = "latex", 
               booktabs = TRUE, 
               linesep = "", 
               col.names = linebreak(tabS1_labs), 
               escape = FALSE, 
               align = "l") %>% 
  kable_styling(latex_options = c("scale_down", 
                                  "HOLD_position")) %>% 
  add_header_above(c(" " = 6, "Particle diameter (nm)" = 3)) %>% 
  pack_rows("DAY286 yeast", 1, 3) %>% 
  pack_rows("ATCC90028", 4, 6) %>% 
  pack_rows("ATCC10231", 7, 9) %>% 
  pack_rows("DAY286 biofilm", 10, 14)

@@

>>>>

# load libraries
library(candidaev)
library(tidyNano)
library(limma)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(kableExtra)

# define colour palette to distinguish between strains
# colours = yellow, blue, green, pink
# levels = yeast, ATCC90028, ATCC10231, biofilm
st_pal <- c("#E69F00", "#56B4E9", "#009E73", "#CC79A7")
```

@@

```{r tableS1Data}
# construct data.frame with nta experiment info
nta_metadata <- data.frame(id = c("ATCC10231_1", "ATCC10231_2", "ATCC10231_3", 
                                  "ATCC90028_1", "ATCC90028_2", "ATCC90028_3", 
                                  "Biofilm_1", "Biofilm_2", "Biofilm_3", 
                                  "Biofilm_4", "Biofilm_5", 
                                  "DAY286_1", "DAY286_2", "DAY286_3"), 
                           strain = c(rep("ATCC10231", 3), 
                                      rep("ATCC90028", 3), 
                                      rep("Biofilm", 5), 
                                      rep("DAY286", 3)), 
                           replicate = c(seq(1, 3), seq(1, 3), 
                                         seq(1, 5), seq(1, 3)), 
                           OD600 = c(0.455*20, 0.557*20, 0.615*20, 
                                     0.310*50, 0.337*50, 0.297*50, 
                                     rep(NA, 5), 
                                     36.2, 39.7, 34.0),
                           culture_vol = c(rep(150, 3),
                                           rep(150, 3), 
                                           rep(1000, 5), 
                                           rep(300, 3)), 
                           ev_vol = c(100, 100, 100, 
                                      100, 100, 100, 
                                      100, 100, 100, 100, 100, 
                                      50, 50, 50), 
                           protein_conc = c(0.192, 0.282, 0.287, 
                                            0.424, 0.716, 0.584, 
                                            0.204, 0.198, 0.297, 0.193, 0.416, 
                                            0.437, 0.360, 0.447), 
                           df = c(250, 750, 750, 
                                  1500, 3000, 3000, 
                                  3000, 2000, 2000, 3000, 6000, 
                                  1500, 1500, 1500), 
                           mean = c(248.5, 218.7, 217.9, 
                                    197.1, 233.1, 191.0, 
                                    119.4, 142.1, 143.3, 133.3, 120.9, 
                                    182.8, 174.5, 164.5), 
                           mode = c(171.9, 155.5, 140.2, 
                                    135.4, 156.7, 138.0, 
                                    100.5, 94.7, 103.1, 103.1, 97.5, 
                                    106.4, 99.2, 100.2), 
                           sd = c(120.1, 112.1, 110.0, 
                                  87.1, 90.7, 68.4, 
                                  47.4, 64.3, 70.2, 60.7, 47.4, 
                                  93.4, 93.9, 83.1), 
                           d10 = c(139.1, 110.2, 118.1, 
                                   122.4, 141.9, 126.0, 
                                   90.2, 88.8, 96.4, 93.2, 89.7, 
                                   97.8, 93.8, 93.0), 
                           d50 = c(211.9, 188.2, 180.3, 
                                   169.4, 209.3, 170.7, 
                                   105.6, 115.0, 110.9, 108.1, 105.9, 
                                   152.8, 147.3, 136.2),
                           d90 = c(393.0, 354, 354.4, 
                                   301.4, 381.1, 285.8, 
                                   151.5, 239.7, 240.8, 210.2, 171.2, 
                                   320.4, 308.0, 290.9),
                           sort_order = c(rep(3, 3), rep(2, 3), 
                                          rep(4, 5), rep(1, 3)), 
                           stringsAsFactors = FALSE) %>% 
  mutate(protein_yield = ev_vol * protein_conc) %>% 
  select(id, strain, replicate, OD600, culture_vol, ev_vol, 
         protein_yield, everything())

# determine the total particle concentration for each biological replicate
nta_total <- nta_techavg %>% 
  nanocount(Strain, Bio_rep, Dilution, 
            param_var = Tech_rep_mean) %>% 
  tidyr::unite("id", Strain, Bio_rep) %>% 
  inner_join(., nta_metadata, by = "id") %>% 
  arrange(sort_order) %>% 
  select(strain, replicate, OD600, culture_vol, ev_vol, protein_yield, 
         protein_conc, Particle_count, mean, mode, sd) %>% 
  mutate(OD600 = case_when(is.na(OD600) == TRUE ~ "NA", 
                           TRUE ~ as.character(OD600)), 
         Particle_count = signif(Particle_count, digits = 3))
```

@@

# construct a table of metadata for each EV biological replicate
tabS1 <- nta_total %>% 
  select(-strain, -ev_vol) %>% 
  mutate_all(linebreak) %>% 
  knitr::kable(caption = tabS1_cap, 
               format = "latex", 
               booktabs = TRUE, 
               linesep = "", 
               col.names = linebreak(tabS1_labs), 
               escape = FALSE, 
               align = "l") %>% 
  kable_styling(latex_options = c("scale_down", 
                                  "HOLD_position")) %>% 
  add_header_above(c(" " = 6, "Particle diameter (nm)" = 3)) %>% 
  pack_rows("DAY286 yeast", 1, 3) %>% 
  pack_rows("ATCC90028", 4, 6) %>% 
  pack_rows("ATCC10231", 7, 9) %>% 
  pack_rows("DAY286 biofilm", 10, 14)

@@

```{r figureS3data}
## Compare mode size of EVs from each strain
# Use ANOVA for comparison
mode_aov <- aov(mode ~ strain, data = nta_total)

# Check assumptions of normality and homogeneous variances
# Plot residuals vs fitted values. Variances seem homogeneous
# plot(mode_aov, 1)

# Check assumption of normality with Shapiro-Wilk test
# Non-significant result indicates data is normal
# shapiro.test(x = residuals(object = mode_aov))

# Check ANOVA is significant (it is) and do TukeyHSD post-hoc
mode_result <- mode_aov %>% 
  tukey_hsd()

# Plot mode size comparison
fs3a <- nta_total %>% 
  ggboxplot(x = "strain", y = "mode", 
            color = "strain", palette = st_pal, add = "jitter", 
            add.params = list(color = "black"), 
            shape = 15, 
            ylab = "Mode particle size (nm)") + 
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 210)) + 
  theme(axis.title.x = element_blank(), 
        legend.position = "none") + 
  stat_compare_means(method = "anova", 
                     label.x = 0.7, 
                     label.y = 205) + 
  stat_pvalue_manual(mode_result[c(2:5), ], label = "p.adj", 
                     y.position = c(180, 190, 205, 170)) + 
  coord_cartesian(clip = "off")

## Compare EV concentration from each strain
# Use ANOVA for comparison
conc_aov <- aov(mode ~ strain, data = nta_total)

# Check assumptions of normality and homogeneous variances
# Plot residuals vs fitted values. Variances seem homogeneous
# plot(conc_aov, 1)

# Check assumption of normality with Shapiro-Wilk test
# Data is non-normal. Therefore use Kruskall-Wallis instead.
# shapiro.test(x = residuals(object = mode_aov))

# Check Kruskall-Wallis is significant (it is) and do Dunn's test post-hoc
conc_kruskall <- kruskal.test(Particle_count ~ strain, data = nta_total)
conc_dunns <- FSA::dunnTest(Particle_count ~ strain, data = nta_total, 
                            method = "bh")[[2]] %>% 
  tidyr::separate(col = Comparison, 
                  into = c("group1", "group2"), 
                  sep = " - ") %>% 
  mutate(p.adj = round(P.adj, digits = 3))

fs3b <- nta_total %>% 
  ggboxplot(x = "strain", y = "Particle_count", 
            color = "strain", palette = st_pal, add = "jitter", 
            add.params = list(color = "black"), 
            shape = 15, 
            ylab = "Total particle concentration / mL") + 
  scale_y_continuous(expand = c(0, 0), 
                     limits = c(0, 1e13)) + 
  theme(axis.title.x = element_blank(), 
        legend.position = "none") + 
  stat_compare_means(method = "kruskal.test", 
                     label.x = 0.82, 
                     label.y = 9.7e12) +
  stat_pvalue_manual(conc_dunns[c(2), ], label = "p.adj", 
                     y.position = 8.8e12) + 
  coord_cartesian(clip = "off")

# arguments for figure S3 export
fS3_args <- list(tiff = list(type = "cairo", compression = "lzw"), 
                postscript = list(bg = "white", fonts = "sans"))
```

```{r figureS3, fig.cap=fS3_cap, fig.height=7, fig.width=5.5, fig.path='figures/', dev=c('pdf', 'postscript', 'tiff'), dev.args=fS3_args, dpi=600}
fS3_cap <- c("\\textbf{MAIN} Sub.")

ggarrange(fs3a, fs3b, 
          labels = c("A", "B"), 
          nrow = 2, 
          ncol = 1, 
          heights = c(1,1), 
          align = "v")
```
